How has the Brazilian Amazon been constructed as a problem? The economy, the environment, the people, and the nation in the presidential speeches since 1985

### 2022-04-11 Draft_April_2022/Full_draft_20220401.rmd
#somewhat nicer tables in pdf #problems with plots, though

---
title: How has the Brazilian Amazon been constructed as a problem? An analysis of
  presidential speeches since 1985
author:
- Livio Silva-Muller^[Phd Candidate, The Geneva Graduate Institute , livio.silva@graduateinstitute.ch]
- Henrique Sposito^[Phd Candidate, The Geneva Graduate Institute Geneva, henrique.sposito@graduateinstitute.ch]
date: "April 2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document:
    reference_docx: template.docx
bibliography: references.bib
csl: chicago.csl
abstract: This paper ...
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(dplyr)
library(ggplot2)
library(tidyr)
library(skimr)
library(scales)
library(RColorBrewer)
library(kableExtra)
library(nnet)
library(stargazer)
Sys.setlocale("LC_ALL", "C")
options(stringsAsFactors = FALSE, scipen = 999)
```

@@

@@

codebook <- tibble::tibble(Problem_Construction = c("National Sovereignty", "Economic Integration",
                                                    "Social Development", "Environmental Conservation"),
                           Description = c("This code constructs the Amazon region and/or forest as an issue of national sovereignty. We understand claims of sovereignty as a particular problem-construction that touches on imaginaries of external threats to territory. Relatedly, we also understand sovereignty as raising concerns about wrong perspectives and criticism from foreign and non-state actors about governments’ actions related to the Brazilian Amazon. In all, it advances the view that the Amazon is Brazilian, foreign, and non-state presence in the region needs to be monitored closely.",
                                           "This code constructs the Amazon region and/or forest as an issue of economic integration. It advances the view that the Amazon needs to be developed and connected to the national economy. This includes expanding the agricultural frontier through incentives, creating a diverse set of infrastructure (roads, dams, internet, radio, energy), fostering differing industries (tourism, mining, cattle, agriculture and so on) through tax-free zones, as well as facilitating the exploitation of natural resources for developmental purposes.",
                                           "This code constructs the Amazon region and/or forest as an issue of social development. It advances the view that Amazon is full of citizens who should have their rights guaranteed. This refers to the construction of schools and universities (right to education), of hospitals (right to health), and of housing (right to house). This also includes guarantees of a dignified life with decent employment, access to water and sanitation, as well as access to electricity, internet, radio, and light. Finally, this includes referrals to culture and the right to vote.",
                                           "This code constructs the Amazon region and/or forest as an issue of conservation. This problem-construction focuses on the value of a standing forest and of the preserved ecosystem in the region. The conservationist narrative advances the view that Amazon should be preserved, deforestation should be halted, and the practices of indigenous and traditional populations should be maintained and fostered. It advances the view that the emission of greenhouse gasses should be halted, that renewable energy should be supported, and that protected areas should be created."),
                           Example = c("Congressman, the fundamental link for Brazil to really head in the direction to prosperity. I would like first, Hu Chunhua,  to thank you for the words of your ambassador to Brazil recognizing our sovereignty over the Amazonian region during that recent episode in the G7 meeting. I would like to thank the Chinese government. For us, this type of public acknowledgement is priceless in your words about this region that is so important to the world and to Brazil. (Bolsonaro 25/10/2019)",

@@

knitr::kable(codebook, format = "simple", caption = "Table 1 - Amazonian Problem-Construction Codebook")

@@

@@

Figure 5, below, illustrates the share of pure-type problem construction by locations in Brazil. We divide locations into Amazonian states (i.e. all the Brazilian states in which the Amazon biome is present), non-Amazonian states (other brazilian states), Brasilia (the capital city in Brazil), and international (countries outside of Brazil). Notice that, consistently, presidents construct the Amazon as an issue of pure economic integration within Amazonian states more often than all other constructions in time. Alternatively, we see that pure conservation becomes increasingly less frequent in time for Amazonian states. As well, pure social development constructions, which were fairly frequent from the early-2000s and onwards, disappeared around 2015 after the mid-2010s in Amazonian states. Nonetheless, we see a large increase from the mid-2000s to the mid-2010s in the frequency in which presidents constructed the Amazon as an issue of pure social development in Brasilia and non-Amazonian states. In Brasilia, also, pure conservation constructions went down drastically during the 1990s, but increased steadily from the mid-2000s onwards. Interestingly, presidents most often constructed the Amazon as an issue of pure environmental conservation internationally only from the mid 2000s until the mid-2010s. Moreover, pure sovereignty constructions did appear in international settings until the mid-2000s.

@@

@@

```{r multimonial model, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
adr <- data.frame(year = c(1985, 1986, 1987), def_year = c(21.050, 21.050, 21.050))
amazon_def_year <- rbind(adr, amazon_def_year)
ama_model <- dplyr::left_join(ama_mx, amazon_def_year, by = "year") # merge
ama_model <- dplyr::left_join(ama_model, AAI, by = "year") # merge
ama_model <- filter(ama_model, location_cat != "Non Identified", mx_cat != "Other Mixed-types") %>% 
  mutate(mx_cat = factor(mx_cat, levels = c("Pure Conservation", "Pure Economic Integration",
                                            "Pure Social Development", "Pure Sovereignty", "Economic Conservation",
                                            "Economic and Social Development", "Sovereign Conservation",
                                            "Sustainable Development","Sovereign Integration",
                                            "Social Conservation", "Other")),
         election_year = ifelse(grepl("1989|1994|1998|2002|2006|2010|2014|2018", year), 1, 0))
# model
model <- multinom(mx_cat ~ location_cat + election_year + def_year + AAI, data = ama_model, model = TRUE, trace = FALSE)
# summary statistics
model_summary <- summary(model)
p_values <- round((1 - pnorm(abs(summary(model)$coefficients/summary(model)$standard.errors)))*2, 5)
# post estimation tests
chi_square <- chisq.test(ama_model$mx_cat, predict(model))
# too much information... Let's just focus on what we want here...
md <- data.frame("Dependent Variables" = c("Economic Integration vs. Environmental Conservation",
                                           "Social Development vs. Environmental Conservation",
                                           "Sovereignty vs. Environmental Conservation"),
                 "Brasilia vs. Amazonian States" = c("-1.024***","-1.002***", "-0.206"),
                 "International vs. Amazonian States" = c("-0.612***", "-1.583***", "-1.427***"),
                 "Non-AM States vs. Amazonian States" = c("0.023", "-0.041", "0.324"),
                 "Election Year" = c("-0.241", "0.028", "-0.684**"),
                 "Annual Deforestation" = c("0.063***", "-0.031*", "0.004"),
                 "Average Inflation" = c("-0.001***", "-0.001***", "-0.0001"))
kableExtra::kable(t(md), format = "simple", caption = "Table 2 - Amazon Problem-Construction by Location")
```

@@

@@

@putnam1988 seminal article on the two level game between domestic and international politics analyzes how negotiations at both levels make policies possible. The author argues that both domestic and international levels need to be taken into consideration when analyzing how, often, domestic politics became entangled in international negotiations. The Amazon, as a region and a forest, has been the topic of international negotiations, national debates, and local policy implementation; though how the Amazon has been constructed as an issue differs on each of these three levels. While in Brasilia or internationally different presidents might construct the Amazon as an issue of conservation, the same presidents might construct the Amazon as an issue of economic integration or social development at the local level. This implies not only that audiences’ priorities in each setting change, but that which policies are appropriate to solve the “Amazon issue” differ. The three level game entails that conservation might be a desirable construction when speaking internationally about the Amazon, but not for local electorates. Such might help explain diverse local implementation gaps between public policy negotiated outside and implemented within the Amazon as perceptions and expectations about the issues the same policy addresses might differ [@alesina2009; @lopez2020].

Whereas presidential discourses at the top matter to define and justify public policy [@zarefsky2004], presidents shape their discourses according to who their audience might be and what "they want to hear". This democratic game contributes to the advancement of ambiguous public policies thought from the outside to the Amazon or contradictory policies from Amazon to the outside. Take, for example, the rural credit offered to local agricultural producers in Amazonian states went from 500 million reais in 1999 to over 4 billion by 2012 [@capobianco2021]. During the same period, the money spent in fighting deforestation in the Amazonian states also increased from … in … to … in … These policies match diverse local, national, and international expectations and match solutions to different problem constructions presented at each of these levels.

@@

We start by cleaning the dataset by removing all accents, non-standard Latin characters, and extra spaces. We then identify all speeches about the Amazon as a region, peoples, or forest out of the 6130 speeches. Our dictionary is composed of a single lexicon stem: “amazon”. In Portuguese, it captures terms such as “Amazonia”, “Amazonica”, “Amazonidas”, “Amazonense(s)”, “Amazonas”, among others. We chose to do so knowing that a few speeches about the Amazon might not contain the lexicon “Amazon”, for example, when the president says “the forest” or uses a pronoun, he or she could be referring to the Amazon. Also, it is likely that whenever a president mentions “deforestation” or derived vocabulary, he or she is speaking about the Amazon. However, as we cannot be sure all the time in those instances (i.e. mentions of “the forest” or “deforestation” could be referring to the Atlantic Forest or the Cerrado). We opt for the safer side and make sure speeches we gather are indeed about the Amazon as a region, forest, or people by only capturing “amazon” related lexicon. We find that 946 speeches are, at least partially, about the Amazon from the 6130.

Using the poldis R package, we extract two sentences before and two sentences after the sentence in which the stem “amazon” appears. By doing so we create our unit of analysis: an Amazonian statement. We use Amazon statements as our unit of analysis for three main reasons. First, presidential speeches can be long and cover a wide variety of topics. For example, the average word count for speeches that mention the stem “amazon” are 2355 words. Working at the level of statements allows us to identify only passages that are meaningful for our specific purpose. A single speech can have more than one Amazonian statement and this statement might be embedded within a different part of the speech. Second, it increases the number of our observations and its variety meaningfully, allowing for more specificity in our analysis. Finally, other authors adopting similar strategies, opt for picking an arbitrary number of words before and after the mention (Kentikellenis and Voeten, 2021). We opt for picking two sentences before and two sentences after, rather than words because a sentence, because sentences usually contain a cohesive idea. Opting for a word threshold, might cut a sentence in the middle and yield an incomplete idea. Nevertheless, this process yields 2014 unique Amazonian statements across the 946 speeches about the Amazon identified previously. Notice that an Amazonian statement can contain two or more matches of the stem “amazon” (i.e. two consecutive sentences in which one of the words covered by the stem appears). In this case, we get two sentences before the first match and two sentences after the last match. On average, an Amazonian statement contains 123 words.

@@

We proceed to operationalize how the Amazon is constructed as a problem. To do so, we develop and apply a qualitative codebook to Amazonian statements. We do so for two reasons. First, and foremost, we believe that a dictionary and other “bag of words” methods might not capture nuances in the texts. These nuances include Amazonian statements that negate certain aspects about the Amazon or statements in which the matched words might not be not directed at the Amazonian region, peoples, or forest. Second, we chose to look at the statements and inductively develop a codebook to verify if the insights from the literature hold within presidential speeches and, if not, add or remove codes, as well as discuss which statements should, and not, be included in each code and why.

After revising the literature (section 2), we went through a random sample of the speeches to develop the codebook. The three policy-cycles (sovereignty, economic integration, and conservation) were clearly identifiable in the speeches. However, once the coding began to take place, however, we noticed two things. One, that some of the Amazon statements were not about Amazon per se. That is, our strategy to identify Amazonian statements yielded a few false positives. These are statements that mention the stem “amazon”, but are not about the Brazilian Amazon region, people, or forest. For example, the president might greet the Governor of the Amazonas in a speech or make a reference to the Venezuelan Amazon. We coded these statements as false positives. Two, we noticed speeches often refer to social development using a different vocabulary of policies and objectives. Those references often differ from references to conservation or economic integration. Hence, we created the code category ‘social development’ as explained below. We consider this to be a finding, but rather than emerging from our analysis, it emerged from codebook development. contain a full description of each code category and the coding practices that inform our procedure.

Table 1 – Qualitative Codebook

Sovereignty
Description: This code constructs the Amazon region and/or forest as an issue of national sovereignty. We understand claims of sovereignty as a particular narrative that touches on imaginaries of external threats to territory. Relatedly, we also understand sovereignty as a particular narrative that raises concerns about wrong perspectives and criticism from foreign and non-state actors about governments’ actions related to the Brazilian Amazon. In all, it advances the view that the Amazon is Brazilian and foreign presence in the region needs to be monitored closely.

Coding practice: we apply this code to references to sovereignty that meaningfully reflect the description above. Simple mentions to the Brazilian Armed Forces, frontiers, or military technology (e.g. SIPAM-SIVAM), are not automatically, by themselves, understood as constructing the Amazon as a problem of sovereignty. When presidents question the authority or interests of someone’s arguments about Amazon, for example, we apply this code as it raises concerns about wrong perspective and criticism. When presidents claim external actors want to internationalize the Amazon territory, we apply this code as it touches upon external threats to territory.

Example: “Senhores deputados. Elo fundamental para que o Brasil realmente rume em direção à prosperidade. Queria primeiro dizer, senhor Hu Chunhua,  eu quero agradecer as palavras do seu embaixador no Brasil reconhecendo a nossa soberania sobre a região Amazônica, no episódio ocorrido há pouco, por ocasião do encontro do G7. Muito obrigado ao governo chinês. Para nós, não têm preço esse reconhecimento público e suas palavras sobre essa região tão importante para o mundo e para o Brasil.” (Bolsonaro 25/10/2019)
Economic Integration
Description: This code constructs the Amazon region and/or forest as an issue of economic integration. It advances the view that the Amazon needs to be developed and connected to the national economy. This includes expanding the agricultural frontier through incentives, creating a diverse set of infrastructure (roads, dams, internet, radio, energy), fostering differing industries (tourism, mining, cattle, agriculture and so on) through tax-free zones, as well as facilitating the exploitation of natural resources for developmental purposes.

Coding practice: We apply this code to references to economic integration that meaningfully reflect the description above. Simple mentions of development, integration, investment, and technology are not automatically coded as economic integration. However, direct mentions of the words “economic development” are coded here. We avoid overlap between this code and social development by considering integration references as targeting infrastructure rather than peoples’ rights. For example, developing infrastructure for internet, radio, and energy are coded as economic integration. However, if the presidents speak about people’s access and rights to information (internet and radio) and electricity, we code as social development. Relatedly, improving macro-structural conditions of the job market is coded as economic integration,, while people’s right to a dignified job is coded as social development. Importantly, we make the analytical choice to include micro-finance instruments as credits and loans in economic development, while grants or donations fall within social development.

Example: "Se me permitem, na Amazônia -que durante tanto tempo ficou adormecida, por falta de uma ação integrada- nós temos já algumas ações muito estruturadoras. Nós, na Amazônia, estamos fazendo uma ligação que passa por Manaus, passa por Boa Vista, Caracaraí, até chegar lá em cima, que é aquela lista vermelha, que vai lá para cima, na direção da Venezuela,é a estrada 174. Essa estrada dará possibilidade da produção da Zona Franca de Manaus ser competitiva, não aqui dentro, mas lá fora mesmo, como é a vocação natural da Zona Franca, exporte, e poderá exportar pelo Caribe.” (FHC 2/7/1997)
Social Development
Description: This code constructs the Amazon region and/or forest as an issue of social development. It advances the view that Amazon is full of citizens who should have their rights guaranteed. This refers to the construction of schools and universities (right to education), of hospitals (right to health), and of housing (right to house). This also includes guarantees of a dignified life with decent employment, access to water and sanitation, as well as access to electricity, internet, radio, and light. Finally, this includes referrals to culture and the right to vote.

Coding practice: We apply this code to references to social development that meaningfully capture the description above. Simple mentions to development are not automatically coded as social development. However, direct mentions supporting “social development” are coded here. An illustrative example relates to speeches outlining the role of the armed forces in providing health services for faraway populations, which we code as social development but not sovereignty.

Example: “Mas como o Estado não tem que ter lucro, o Estado precisa garantir a cidadania, nós achamos que o cidadão que mora às margens do rio Amazonas, a 600 Km de Manaus, ele tem que ter direito a ter luz na sua casa, a ter geladeira, a ter televisão e a ver sua novela. Já investimos 14 bilhões de reais nesse programa, em três anos e meio. Sabe quantos postes nós já colocamos? Um milhão de quilômetros de fio.” (Lula 20/11/2009)
Conservation
Description: This code constructs the Amazon region and/or forest as an issue of conservation. This narrative focuses on the value of a standing forest and of the preserved ecosystem in the region. The conservationist narrative advances the view that Amazon should be preserved, deforestation should be halted, and the practices of indigenous peoples should be maintained and fostered. It advances the view that the emission of greenhouse gasses should be halted, that renewable energy should be supported, and that protected areas should be created.

@@

Example: “Adotei medidas de caráter emergencial, suspendi a exportação de madeiras em toras, suspendi incentivos fiscais e creditícios na Amazônia a projetos que podiam causar danos ambientais e instituí a obrigatoriedade do licenciamento para a atividade de extração de ouro com a proibição da utilização do mercúrio. Dando início à reestruturação do sistema governamental de controle e preservação do meio ambiente, criei o Instituto Brasileiro do Meio Ambiente e Recursos Naturais, IBAMA, que está entregue à capacidade do Dr. Fernando César Mesquita.” (Sarney, 20/7/1989)
False positive
Description: statements that were assigned as Amazonian but are about the Amazon.

>>>>

---
title: How has the Brazilian Amazon been constructed as a problem? The economy, the environment, the people, and the nation in the presidential speeches since 1985
author:
- Livio Silva-Muller^[Phd Candidate, The Geneva Graduate Institute , livio.silva@graduateinstitute.ch]
- Henrique Sposito^[Phd Candidate, The Geneva Graduate Institute Geneva, henrique.sposito@graduateinstitute.ch]
date: "April 2022"
output:
  html_document:
    df_print: paged
  word_document:
    reference_docx: template.docx
  pdf_document: default
bibliography: references.bib
csl: chicago.csl
abstract: This paper ...
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(dplyr)
library(ggplot2)
library(tidyr)
library(skimr)
library(scales)
library(RColorBrewer)
library(kableExtra)
library(nnet)
library(stargazer)
library(kableExtra)
Sys.setlocale("LC_ALL", "C")
options(stringsAsFactors = FALSE, scipen = 999)
```

@@

\begin{landscape}

@@

codebook <- tibble::tibble("Problem Construction" = c("National Sovereignty", "Economic Integration","Social Development", "Environmental Conservation"),
                           Description = c("This code constructs the Amazon region and/or forest as an issue of national sovereignty. We understand claims of sovereignty as a particular problem-construction that touches on imaginaries of external threats to territory. Relatedly, we also understand sovereignty as raising concerns about wrong perspectives and criticism from foreign and non-state actors about governments’ actions related to the Brazilian Amazon. In all, it advances the view that the Amazon is Brazilian, foreign, and non-state presence in the region needs to be monitored closely.","This code constructs the Amazon region and/or forest as an issue of economic integration. It advances the view that the Amazon needs to be developed and connected to the national economy. This includes expanding the agricultural frontier through incentives, creating a diverse set of infrastructure (roads, dams, internet, radio, energy), fostering differing industries (tourism, mining, cattle, agriculture and so on) through tax-free zones, as well as facilitating the exploitation of natural resources for developmental purposes.", "This code constructs the Amazon region and/or forest as an issue of social development. It advances the view that Amazon is full of citizens who should have their rights guaranteed. This refers to the construction of schools and universities (right to education), of hospitals (right to health), and of housing (right to house). This also includes guarantees of a dignified life with decent employment, access to water and sanitation, as well as access to electricity, internet, radio, and light. Finally, this includes referrals to culture and the right to vote.", "This code constructs the Amazon region and/or forest as an issue of conservation. This problem-construction focuses on the value of a standing forest and of the preserved ecosystem in the region. The conservationist narrative advances the view that Amazon should be preserved, deforestation should be halted, and the practices of indigenous and traditional populations should be maintained and fostered. It advances the view that the emission of greenhouse gasses should be halted, that renewable energy should be supported, and that protected areas should be created."), Example = c("Congressman, the fundamental link for Brazil to really head in the direction to prosperity. I would like first, Hu Chunhua,  to thank you for the words of your ambassador to Brazil recognizing our sovereignty over the Amazonian region during that recent episode in the G7 meeting. I would like to thank the Chinese government. For us, this type of public acknowledgement is priceless in your words about this region that is so important to the world and to Brazil. (Bolsonaro 25/10/2019)",

@@

kbl(codebook, booktabs = T,caption = "Table 1 - Amazonian Problem-Construction Codebook")%>% kable_styling(latex_options = c("striped", "repeat_header", "scale_down"), position= "center", font_size = 9, full_width = TRUE) %>%
  column_spec(1, width="3cm", bold=TRUE)%>%
  column_spec(2, width="9cm") %>%
  column_spec(3, width = "9cm")

@@

\end{landscape}

@@

@@

Figure 5, above, illustrates the share of pure-type problem construction by locations in Brazil. We divide locations into Amazonian states (i.e. all the Brazilian states in which the Amazon biome is present), non-Amazonian states (other brazilian states), Brasilia (the capital city in Brazil), and international (countries outside of Brazil). Notice that, consistently, presidents construct the Amazon as an issue of pure economic integration within Amazonian states more often than all other constructions in time. Alternatively, we see that pure conservation becomes increasingly less frequent in time for Amazonian states. As well, pure social development constructions, which were fairly frequent from the early-2000s and onwards, disappeared around 2015 after the mid-2010s in Amazonian states. Nonetheless, we see a large increase from the mid-2000s to the mid-2010s in the frequency in which presidents constructed the Amazon as an issue of pure social development in Brasilia and non-Amazonian states. In Brasilia, also, pure conservation constructions went down drastically during the 1990s, but increased steadily from the mid-2000s onwards. Interestingly, presidents most often constructed the Amazon as an issue of pure environmental conservation internationally only from the mid 2000s until the mid-2010s. Moreover, pure sovereignty constructions did appear in international settings until the mid-2000s.

@@

@@

```{r multimonial model, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
adr <- data.frame(year = c(1985, 1986, 1987), def_year = c(21.050, 21.050, 21.050))
amazon_def_year <- rbind(adr, amazon_def_year)
ama_model <- dplyr::left_join(ama_mx, amazon_def_year, by = "year") # merge
ama_model <- dplyr::left_join(ama_model, AAI, by = "year") # merge
ama_model <- filter(ama_model, location_cat != "Non Identified", mx_cat != "Other Mixed-types") %>% 
  mutate(mx_cat = factor(mx_cat, levels = c("Pure Conservation", "Pure Economic Integration",
                                            "Pure Social Development", "Pure Sovereignty", "Economic Conservation",
                                            "Economic and Social Development", "Sovereign Conservation",
                                            "Sustainable Development","Sovereign Integration",
                                            "Social Conservation", "Other")),
         election_year = ifelse(grepl("1989|1994|1998|2002|2006|2010|2014|2018", year), 1, 0))
# model
model <- multinom(mx_cat ~ location_cat + election_year + def_year + AAI, data = ama_model, model = TRUE, trace = FALSE)
# summary statistics
model_summary <- summary(model)
p_values <- round((1 - pnorm(abs(summary(model)$coefficients/summary(model)$standard.errors)))*2, 5)
# post estimation tests
chi_square <- chisq.test(ama_model$mx_cat, predict(model))
# too much information... Let's just focus on what we want here...

md <- tibble("Dependent Variable" = c("Economic Integration vs Environmental Conservation",
                                           "Social Development vs Environmental Conservation",
                                           "Sovereignty vs Environmental Conservation"),
                 "Brasilia vs Amazonian States" = c("-1.024***","-1.002***", "-0.206"),
                 "International vs Amazonian States" = c("-0.612***", "-1.583***", "-1.427***"),
                 "Other States vs Amazonian States" = c("0.023", "-0.041", "0.324"),
                 "Election Year" = c("-0.241", "0.028", "-0.684**"),
                 "Annual Deforestation" = c("0.063***", "-0.031*", "0.004"),
                 "Average Inflation" = c("-0.001***", "-0.001***", "-0.0001"))

kbl(t(md), booktabs = T,caption = "Table 2 - Amazon Problem-Construction by Location", align = "c")%>% kable_styling(latex_options = c("striped", "scale_down"), position= "center", font_size = 9, full_width = TRUE) %>%
  column_spec(1, width="5 cm", bold=TRUE, latex_valign = "m")%>%
  column_spec(2, width="5cm", latex_valign = "m") %>%
  column_spec(3, width = "5cm", latex_valign = "m") %>%
  column_spec(4, width = "5cm", latex_valign = "m")

@@

@putnam1988 seminal article on the two level game between domestic and international politics analyzes how negotiations at both levels make policies possible. The author argues that both domestic and international levels need to be taken into consideration when analyzing how, often, domestic politics became entangled in international negotiations. The Amazon, as a region and a forest, has been the topic of international negotiations, national debates, and local policy implementation; though how the Amazon has been constructed as an issue differs on each of these three levels. While in Brasilia or internationally different presidents might construct the Amazon as an issue of conservation, the same presidents might construct the Amazon as an issue of economic integration or social development at the local level. This implies not only that audiences' priorities in each setting change, but that which policies are appropriate to solve the "Amazon issue" differ. The three level game entails that conservation might be a desirable construction when speaking internationally about the Amazon, but not for local electorates. Such might help explain diverse local implementation gaps between public policy negotiated outside and implemented within the Amazon as perceptions and expectations about the issues the same policy addresses might differ [@alesina2009; @lopez2020].

Whereas presidential discourses at the top matter to define and justify public policy [@zarefsky2004], presidents shape their discourses according to who their audience might be and what "they want to hear". This democratic game contributes to the advancement of ambiguous public policies thought from the outside to the Amazon or contradictory policies from Amazon to the outside. Take, for example, the rural credit offered to local agricultural producers in Amazonian states went from 500 million reais in 1999 to over 4 billion by 2012 [@capobianco2021]. During the same period, the money spent in fighting deforestation in the Amazonian states also increased from ... in ... to ... in ... These policies match diverse local, national, and international expectations and match solutions to different problem constructions presented at each of these levels.

@@

We start by cleaning the dataset by removing all accents, non-standard Latin characters, and extra spaces. We then identify all speeches about the Amazon as a region, peoples, or forest out of the 6130 speeches. Our dictionary is composed of a single lexicon stem: "amazon". In Portuguese, it captures terms such as "Amazonia", "Amazonica", "Amazonidas", "Amazonense(s)", "Amazonas", among others. We chose to do so knowing that a few speeches about the Amazon might not contain the lexicon "Amazon", for example, when the president says "the forest" or uses a pronoun, he or she could be referring to the Amazon. Also, it is likely that whenever a president mentions "deforestation" or derived vocabulary, he or she is speaking about the Amazon. However, as we cannot be sure all the time in those instances (i.e. mentions of "the forest" or "deforestation" could be referring to the Atlantic Forest or the Cerrado). We opt for the safer side and make sure speeches we gather are indeed about the Amazon as a region, forest, or people by only capturing "amazon" related lexicon. We find that 946 speeches are, at least partially, about the Amazon from the 6130.

Using the poldis R package, we extract two sentences before and two sentences after the sentence in which the stem "amazon" appears. By doing so we create our unit of analysis: an Amazonian statement. We use Amazon statements as our unit of analysis for three main reasons. First, presidential speeches can be long and cover a wide variety of topics. For example, the average word count for speeches that mention the stem "amazon" are 2355 words. Working at the level of statements allows us to identify only passages that are meaningful for our specific purpose. A single speech can have more than one Amazonian statement and this statement might be embedded within a different part of the speech. Second, it increases the number of our observations and its variety meaningfully, allowing for more specificity in our analysis. Finally, other authors adopting similar strategies, opt for picking an arbitrary number of words before and after the mention (Kentikellenis and Voeten, 2021). We opt for picking two sentences before and two sentences after, rather than words because a sentence, because sentences usually contain a cohesive idea. Opting for a word threshold, might cut a sentence in the middle and yield an incomplete idea. Nevertheless, this process yields 2014 unique Amazonian statements across the 946 speeches about the Amazon identified previously. Notice that an Amazonian statement can contain two or more matches of the stem "amazon" (i.e. two consecutive sentences in which one of the words covered by the stem appears). In this case, we get two sentences before the first match and two sentences after the last match. On average, an Amazonian statement contains 123 words.

@@

We proceed to operationalize how the Amazon is constructed as a problem. To do so, we develop and apply a qualitative codebook to Amazonian statements. We do so for two reasons. First, and foremost, we believe that a dictionary and other "bag of words" methods might not capture nuances in the texts. These nuances include Amazonian statements that negate certain aspects about the Amazon or statements in which the matched words might not be not directed at the Amazonian region, peoples, or forest. Second, we chose to look at the statements and inductively develop a codebook to verify if the insights from the literature hold within presidential speeches and, if not, add or remove codes, as well as discuss which statements should, and not, be included in each code and why.

After revising the literature (section 2), we went through a random sample of the speeches to develop the codebook. The three policy-cycles (sovereignty, economic integration, and conservation) were clearly identifiable in the speeches. However, once the coding began to take place, however, we noticed two things. One, that some of the Amazon statements were not about Amazon per se. That is, our strategy to identify Amazonian statements yielded a few false positives. These are statements that mention the stem "amazon", but are not about the Brazilian Amazon region, people, or forest. For example, the president might greet the Governor of the Amazonas in a speech or make a reference to the Venezuelan Amazon. We coded these statements as false positives. Two, we noticed speeches often refer to social development using a different vocabulary of policies and objectives. Those references often differ from references to conservation or economic integration. Hence, we created the code category 'social development' as explained below. We consider this to be a finding, but rather than emerging from our analysis, it emerged from codebook development. contain a full description of each code category and the coding practices that inform our procedure.

Table 1 -- Qualitative Codebook

Sovereignty Description: This code constructs the Amazon region and/or forest as an issue of national sovereignty. We understand claims of sovereignty as a particular narrative that touches on imaginaries of external threats to territory. Relatedly, we also understand sovereignty as a particular narrative that raises concerns about wrong perspectives and criticism from foreign and non-state actors about governments' actions related to the Brazilian Amazon. In all, it advances the view that the Amazon is Brazilian and foreign presence in the region needs to be monitored closely.

Coding practice: we apply this code to references to sovereignty that meaningfully reflect the description above. Simple mentions to the Brazilian Armed Forces, frontiers, or military technology (e.g. SIPAM-SIVAM), are not automatically, by themselves, understood as constructing the Amazon as a problem of sovereignty. When presidents question the authority or interests of someone's arguments about Amazon, for example, we apply this code as it raises concerns about wrong perspective and criticism. When presidents claim external actors want to internationalize the Amazon territory, we apply this code as it touches upon external threats to territory.

Example: "Senhores deputados. Elo fundamental para que o Brasil realmente rume em direção à prosperidade. Queria primeiro dizer, senhor Hu Chunhua, eu quero agradecer as palavras do seu embaixador no Brasil reconhecendo a nossa soberania sobre a região Amazônica, no episódio ocorrido há pouco, por ocasião do encontro do G7. Muito obrigado ao governo chinês. Para nós, não têm preço esse reconhecimento público e suas palavras sobre essa região tão importante para o mundo e para o Brasil." (Bolsonaro 25/10/2019) Economic Integration Description: This code constructs the Amazon region and/or forest as an issue of economic integration. It advances the view that the Amazon needs to be developed and connected to the national economy. This includes expanding the agricultural frontier through incentives, creating a diverse set of infrastructure (roads, dams, internet, radio, energy), fostering differing industries (tourism, mining, cattle, agriculture and so on) through tax-free zones, as well as facilitating the exploitation of natural resources for developmental purposes.

Coding practice: We apply this code to references to economic integration that meaningfully reflect the description above. Simple mentions of development, integration, investment, and technology are not automatically coded as economic integration. However, direct mentions of the words "economic development" are coded here. We avoid overlap between this code and social development by considering integration references as targeting infrastructure rather than peoples' rights. For example, developing infrastructure for internet, radio, and energy are coded as economic integration. However, if the presidents speak about people's access and rights to information (internet and radio) and electricity, we code as social development. Relatedly, improving macro-structural conditions of the job market is coded as economic integration,, while people's right to a dignified job is coded as social development. Importantly, we make the analytical choice to include micro-finance instruments as credits and loans in economic development, while grants or donations fall within social development.

Example: "Se me permitem, na Amazônia -que durante tanto tempo ficou adormecida, por falta de uma ação integrada- nós temos já algumas ações muito estruturadoras. Nós, na Amazônia, estamos fazendo uma ligação que passa por Manaus, passa por Boa Vista, Caracaraí, até chegar lá em cima, que é aquela lista vermelha, que vai lá para cima, na direção da Venezuela,é a estrada 174. Essa estrada dará possibilidade da produção da Zona Franca de Manaus ser competitiva, não aqui dentro, mas lá fora mesmo, como é a vocação natural da Zona Franca, exporte, e poderá exportar pelo Caribe." (FHC 2/7/1997) Social Development Description: This code constructs the Amazon region and/or forest as an issue of social development. It advances the view that Amazon is full of citizens who should have their rights guaranteed. This refers to the construction of schools and universities (right to education), of hospitals (right to health), and of housing (right to house). This also includes guarantees of a dignified life with decent employment, access to water and sanitation, as well as access to electricity, internet, radio, and light. Finally, this includes referrals to culture and the right to vote.

Coding practice: We apply this code to references to social development that meaningfully capture the description above. Simple mentions to development are not automatically coded as social development. However, direct mentions supporting "social development" are coded here. An illustrative example relates to speeches outlining the role of the armed forces in providing health services for faraway populations, which we code as social development but not sovereignty.

Example: "Mas como o Estado não tem que ter lucro, o Estado precisa garantir a cidadania, nós achamos que o cidadão que mora às margens do rio Amazonas, a 600 Km de Manaus, ele tem que ter direito a ter luz na sua casa, a ter geladeira, a ter televisão e a ver sua novela. Já investimos 14 bilhões de reais nesse programa, em três anos e meio. Sabe quantos postes nós já colocamos? Um milhão de quilômetros de fio." (Lula 20/11/2009) Conservation Description: This code constructs the Amazon region and/or forest as an issue of conservation. This narrative focuses on the value of a standing forest and of the preserved ecosystem in the region. The conservationist narrative advances the view that Amazon should be preserved, deforestation should be halted, and the practices of indigenous peoples should be maintained and fostered. It advances the view that the emission of greenhouse gasses should be halted, that renewable energy should be supported, and that protected areas should be created.

@@

Example: "Adotei medidas de caráter emergencial, suspendi a exportação de madeiras em toras, suspendi incentivos fiscais e creditícios na Amazônia a projetos que podiam causar danos ambientais e instituí a obrigatoriedade do licenciamento para a atividade de extração de ouro com a proibição da utilização do mercúrio. Dando início à reestruturação do sistema governamental de controle e preservação do meio ambiente, criei o Instituto Brasileiro do Meio Ambiente e Recursos Naturais, IBAMA, que está entregue à capacidade do Dr. Fernando César Mesquita." (Sarney, 20/7/1989) False positive Description: statements that were assigned as Amazonian but are about the Amazon.

Example: "cumprimentar os presidentes de bancos publicos do banco do brasil alexandre abreu do bndes luciano coutinho da caixa miriam belchior do banco do nordeste marcos costa holanda do banco da amazonia valmir pedro rossi o vicepresidente do banco do brasil cesar borges da diretoria de infraestrutura queria cumprimentar as senhoras e os senhores dirigentes de agencias reguladoras dirigir um cumprimento especial ao senhor robson andrade presidente da confederacao nacional da industria por intermedio de quem cumprimento todos os empresarios das federacoes aqui presentes" (Rousseff, 9/7/2015)
